---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
options(digits = 4, width = 120)
```

# nmatch: Fuzzy matching for proper names

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R-CMD-check](https://github.com/epicentre-msf/nmatch/workflows/R-CMD-check/badge.svg)](https://github.com/epicentre-msf/nmatch/actions)
<!-- badges: end -->

Compare or match proper names from different sources, accounting for potential
variation in format, name order, and spelling (e.g. "Beyoncé Knowles" vs.
"KNOWLES-CARTER, Beyonce"). Specifically, uses string-standardization to account
for variation in punctuation, accents, and character case, token-permutation to
account for variation in name order, and fuzzy matching to handle alternate
spellings.

### Installation

Install from GitHub with:

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("epicentre-msf/nmatch")
```

### Example usage

```{r, message=FALSE}
library(nmatch)
library(dplyr)
library(purrr)
library(tidyr)

# example data with names to compare
data("names_ex")

# compare with default settings
names_ex %>% 
  mutate(is_match = nmatch(name_source1, name_source2))

# return full match details
names_ex %>% 
  mutate(match_df = purrr::map2(name_source1, name_source2, nmatch, return_full = TRUE)) %>% 
  tidyr::unnest("match_df")

# use a custom function to classify matches, based on any of the vars returned
# when return_full = TRUE (i.e. k_x, k_y, k_align, n_match, dist_total)
# note that fn must include argument `...` even if not used
classify_matches <- function(k_align, dist_total, ...) {
  k_align >= 1 & dist_total <= 3
}

names_ex %>% 
  mutate(is_match = nmatch(name_source1, name_source2, eval_fn = classify_matches))
```


#### Using with [fuzzyjoin](https://github.com/dgrtwo/fuzzyjoin)

```{r}
library(fuzzyjoin)

# example hospital datasets (IPD and ICU) with patients names in different
# formats (note these names are just random combinations of common French names)
data("dat_ipd")
data("dat_icu")

# join IPD and ICU datasets by name variables, using fuzzy matching with nmatch()
fuzzyjoin::fuzzy_join(
  dat_ipd,
  dat_icu,
  by = c("name_ipd" = "name_icu"),
  match_fun = nmatch::nmatch,
  mode = "left"
)
```

Note that matching large datasets using `fuzzy_join()` and `nmatch()` may be
very slow. The nmatch package will eventually contain stand-alone join functions
that are optimized for name-matching.
